Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
incorporate Abby's stats in intro

*(Abby has offered to update stats in this paragraph:)*

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions (REF)*. (**check if deforestation statistic below includes natural fire, exclude here if it does.**)
In recent years, the remaining CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.
Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).

>>>>

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions (REF)*. (**check if deforestation statistic below includes natural fire, exclude here if it does.**)
In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^−1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.
Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr^-1^ from YEAR-YEAR (Pan et al 2011; **UPDATE**), reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
